---
title: "Food Supply and Death Attributed to Obesity"
author: "Ken Ye"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
# Load libraries
library(knitr)
library(kableExtra)
library(dplyr)
library(glmnet)
library(mice)
library(ggplot2)
library(scatterplot3d)
library(car)
```

```{r include = FALSE}
# Load data set
supply <- read.csv("Data/dietary-composition-by-country.csv")
death <- read.csv("Data/share-of-deaths-obesity.csv")
# gdp <- read.csv("Data/gdp-per-capita-worldbank.csv")
income.group <- read.csv("Data/world-banks-income-groups.csv")
```

# Introduction

With an alarming 41.9% of the U.S. population classified as obese, as revealed
by the 2011 National Health and Nutrition Examination Survey (NHANES) [1], the
implications of this health crisis become undeniably significant. However, the
impact becomes even more pronounced when considered globally, where obesity
contributed to 8% of deaths in 2017 --- an alarming surge from 4.5% in 1990 [2].

Obesity, beyond its visual manifestations, is intricately linked to severe
health conditions, including heart disease, stroke, type 2 diabetes, and various
cancers. These conditions stand as leading causes of preventable, premature
deaths. The economic burden is equally staggering, with the estimated annual
medical cost of obesity in the United States reaching nearly \$173 billion in
2019 dollars. Notably, medical costs for adults grappling with obesity were
\$1,861 higher than those for individuals maintaining a healthy weight [1].

As we delve into the complex web of factors contributing to obesity, the role of
food emerges as a central player. Continuous overconsumption, a potential
precursor to weight gain and obesity, emphasizes the critical importance of
understanding food composition. Certain foods, due to their nutrient content,
can significantly influence overall health outcomes.

However, the narrative extends beyond food alone. Economic status emerges as a
pivotal consideration, acting as a covariate in my analysis. Death rates
attributed to obesity exhibit variations among countries, and a fundamental
differentiator lies in their economic standing. Developed countries, buoyed by
high incomes, may boast advanced medical systems capable of mitigating
obesity-related deaths. Paradoxically, these countries may also grapple with
higher obesity prevalence. On the contrary, developing nations, constrained by
lower incomes, may face a different set of challenges, potentially emphasizing
malnutrition over obesity.

In this analysis, the focal point is discerning the effects of a country's food
supply and economic status on death rates attributed to obesity. The
implications of this study are far-reaching, offering valuable insights for
policymakers eager to curb obesity-related deaths and aiding consumers in
crafting diets resilient to the detrimental effects of obesity.

A dual-stage methodology will be employed. Initially, simple linear regression
sheds light on the potential impact of individual food groups and a country's
economic status. This analysis guides us towards identifying statistically
significant associations. Subsequently, the methodology advances to Lasso
regression, a more sophisticated model aimed at pinpointing the most influential
variables while mitigating the risk of overfitting.

# Data

## Data Sources

Three separate datasets, totaling 26 predictor variables and 1 response
variable, are used in this analysis:

1)  Daily caloric supply (OWID based on UN FAO & historical sources)

Predictor variables (after renaming): `Miscellaneous`, `Alcohol`, `Animal.fat`,
`Vegetable.oils`, `Oilcrops`, `Fish.and.seafood`, `Sugar.crops`,
`Sugar.sweeteners`, `Starchy.roots`, `Meat.other`, `Meat.sheep.and.goat`,
`Meat.pig`, `Meat.poultry`, `Meat.beef`, `Eggs`, `Milk`, `Nuts`, `Fruit`,
`Vegetables`, `Pulses`, `Cereals.other`, `Barley`, `Maize`, `Rice`, `Wheat` ---
each one (25 total) is a food group available for consumption, continuous, unit
(kcal per day per capita)

Source: Our World in Data based on UN FAO & historical sources --- processed by
Our World in Data

Link: <https://www.fao.org/faostat/en/#data/FBS>

2)  World Bank income classification

Predictor variable (after renaming): `Income.group` --- nominal, four categories
(Low-income countries, Lower-middle-income countries, Upper-middle-income
countries, High-income countries)

Source: World Bank (2022) --- processed by Our World in Data

Link:
<https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups>

3)  Share of total deaths that are from all causes attributed to high body-mass
    index, in both sexes aged age-standardized

Response variable (after renaming): `Death.obesity` --- continuous, unit (%)

Source: IHME, Global Burden of Disease Study (2019) --- processed by Our World
in Data

Link: <https://ourworldindata.org/obesity>

## Data Cleaning & Merging

```{r}
# Filter out 2019 data (newest for both data sets)
# Filter out entities without code (not countries)
supply <- supply |>
  filter(Code != "", Year == 2019) |>
  select(-Code, -Year)

death <- death |>
  filter(Code != "",Year == 2019) |>
  select(-Code, -Year)

# gdp <- gdp |>
#  filter(Code != "",Year == 2019) |>
#  select(-Code, -Year)

income.group <- income.group |>
  filter(Code != "",Year == 2019) |>
  select(-Code, -Year)

# Rename variables
colnames(supply) <- c("Country", "Miscellaneous", "Alcohol", "Animal.fat", "Vegetable.oils", "Oilcrops", "Fish.and.seafood", "Sugar.crops", "Sugar.sweeteners", "Starchy.roots", "Meat.other", "Meat.sheep.and.goat", "Meat.pig", "Meat.poultry", "Meat.beef", "Eggs", "Milk", "Nuts", "Fruit", "Vegetables", "Pulses", "Cereals.other", "Barley", "Maize", "Rice", "Wheat")

colnames(death) <- c("Country", "Death.obesity")

# colnames(gdp) <- c("Country", "GDP.per.capita")

colnames(income.group) <- c("Country", "Income.group")
```

```{r}
# Merge data sets
df <- merge(supply, income.group, by = "Country", all = FALSE)
# df <- merge(df, gdp, by = "Country", all = FALSE)
df <- merge(df, death, by = "Country", all = FALSE)
```

To ensure chronological congruence in the analysis, I selected the latest year
(2019) for which data was available in all three datasets and excluded any
observations for entities that are not countries (e.g. Africa), focusing on
individual countries. Renaming variables enhanced legibility. Finally, I merged
the three datasets using the inner join method on the variable `Country`. The
resulting merged dataset, named `df`, comprises 181 observations and 28 columns.

## Missing Data Imputation

```{r include = FALSE}
# Create a mice imputation object
mice_imputation <- mice(df, method = "pmm")

# Perform the imputation
df <- complete(mice_imputation)
```

In my analysis, I encountered missing values in certain food groups across
countries. Removing these entries would significantly reduce the dataset
entries, leaving only 100 out of 174 countries with 45 degrees of freedom for
model fitting (if all columns are utilized for the model). This approach would
not be ideal for accurately representing global patterns or obtaining robust
model estimates.

To address this challenge, I opted for imputation using the "predictive mean
matching"(PMM) method from the `mice` package. PMM ensures that imputed values
are drawn from observed values with similar predicted values, preserving the
distributional characteristics of the data. This method allows us to utilize the
full dataset, providing a more comprehensive and representative view of the
relationships between food supply patterns and obesity death rates across
diverse countries. Imputing missing values with PMM contributes to the validity
of our inferences and enhances the generalizability of my findings to a broader
global context.

## EDA

### Obesity-attributed Death Visualization

![Share of Deaths Attributed to Obesity Across the
Globe](Data/share-of-deaths-obesity.png)

The distribution of obesity-attributed death varies significantly on a global
scale, as illustrated in the above map.

In numerous middle-income nations, notably in Eastern Europe, Central Asia,
North Africa, and Latin America, over 15% of deaths were linked to obesity. This
pattern likely arises from a high prevalence of obesity coupled with
comparatively poorer overall health and healthcare systems, especially when
compared to high-income countries experiencing similar levels of obesity.

Contrastingly, in most high-income countries, this proportion ranges between 8%
to 10%, approximately half of that observed in many middle-income countries.
Notable exceptions among affluent nations are Japan and South Korea, where only
around 5% of premature deaths are attributed to obesity.

Turning to low-income countries, particularly those in Sub-Saharan Africa,
obesity accounts for less than 5% of deaths, underscoring a distinct
epidemiological profile in these regions.

```{r}
df$Income.group <- factor(df$Income.group, levels = c("Low-income countries", "Lower-middle-income countries", "Upper-middle-income countries", "High-income countries"))

ggplot(df, aes(x = Income.group, y = Death.obesity)) +
  geom_boxplot(fill = "skyblue", color = "darkblue", alpha = 0.7) +
  labs(title = "Share of Deaths Attributed to Obesity", x = "Income Group", y = "Obesity-attributed Death Rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"))
```

The graph above, utilizing 2019 data, provides a clearer depiction of how the
obesity-attributed death rate varies across countries' economic statuses.
Notably, as a country progresses through income group classifications from
Low-income to Lower-middle-income to Upper-middle-income, there is an observable
increase in the obesity-attributed death rate. Intriguingly, when a country
transitions from Upper-middle-income to High-income, the obesity-attributed
death rate begins to decrease once more. This pattern adds an intriguing layer
to the analysis, prompting further exploration into the nuanced relationship
between economic status and obesity-related mortality.

### Exploration of Interaction Effects

```{r}
scatterplot3d(df$Rice, df$Pulses, df$Death.obesity, color = df$Pulses,
              main = "3D Scatter Plot of Rice and Pulses Interaction",
              xlab = "Rice", ylab = "Pulses", zlab = "Death Obesity",
              angle = 55, pch = 16, color.palette = heat.colors)
```

```{r results = "hide"}
pulses.rice <- lm(Death.obesity ~ Pulses * Rice, data = df)
summary(pulses.rice)
```

```{r results = "hide"}
poultry.vege <- lm(Death.obesity ~ Meat.poultry * Vegetables, data = df)
summary(poultry.vege)
```

```{r results = "hide"}
beef.starch <- lm(Death.obesity ~ Meat.beef * Starchy.roots, data = df)
summary(beef.starch)
```

During meal consumption, certain food pairings emerge as common occurrences ---
rice (`Rice`) paired with beans (`Pulses`), chicken (`Meat.poultry`) alongside
broccoli (`Vegetables`), and steak (`Meat.beef)` accompanied by mashed potatoes
(`Starchy.roots`), to name a few. Exploring potential interaction effects
between the supply of these food groups could shed light on nuanced dietary
patterns and their impact on obesity-attributed death. However, an examination
of the coefficients for these interaction terms reveals a lack of statistical
significance. Consequently, the exploration of interaction effects is not
pursued further in the subsequent models.

# Methodology

```{r}
# Split data into 80% training, 20% testing
x <- model.matrix(Death.obesity ~ . - Country, df)[,-1]
y <- df$Death.obesity

set.seed(1)
train <- sample(1:nrow(x), floor(nrow(x)*0.8))
test <- setdiff(1:nrow(x), train)
y.test <- y[test]
```

The 181 observations are randomly divided into 80% training and 20% testing
sets.

## Simple Linear Regression Model

```{r results = "hide"}
lm <- lm(y ~ x, subset = train)
lm.pred <- predict(lm, newx = x[test,])
summary(lm)
```

```{r results = "hide"}
vif_data <- data.frame(Variables = colnames(x), VIF = numeric(ncol(x)))

for (i in 1:ncol(x)) {
  lm_i <- lm(x[, i] ~ x[, -i])
  vif_i <- 1 / (1 - summary(lm_i)$r.squared)
  vif_data$VIF[i] <- vif_i
}

print(vif_data)
```

The simple linear regression model is employed to predict the obesity-attributed
death rate based on the supply of 25 food groups and the country's income group.
The model exhibited a Multiple R-squared of 0.6466, indicating a reasonable
level of explained variability. Significant predictors included specific food
groups such as fish and seafood, sheep and goat, pig, poultry, beef, and wheat.
Among income groups, lower-middle-income countries, upper-middle-income
countries, and high-income countries also demonstrated notable effects. The
F-statistic was significant (p-value: 2.181e-15), affirming the model's overall
significance. In addition, while checking for multicollinearity issues, all
variables except for upper-middle-income countries (VIF = 5.4) and high-income
countries (VIF = 8.9) had VIF values \< 5, indicating low or no
multicollinearity issues. This initial analysis provides valuable insights into
factors influencing obesity-related mortality, paving the way for further
refinement and exploration through techniques like Lasso regression.

Lasso regression provides several advantages for refining predictive models. One
key feature is the introduction of a regularization term that encourages
sparsity in coefficient estimates, facilitating automatic variable selection.
This proves particularly beneficial when dealing with a substantial number of
predictors, such as the 26 predictors in this case, where prior knowledge about
their significance may be limited. Lasso tends to shrink certain coefficients to
zero, effectively excluding irrelevant variables and enhancing model
interpretability.

In addition to variable selection, Lasso addresses multicollinearity issues by
favoring a unique subset of correlated predictors. This not only contributes to
model stability but also ensures interpretability by focusing on the most
relevant information. The regularization incorporated in Lasso goes beyond
enhancing model generalization; it also promotes robustness, making the model
more resilient to fluctuations in the data and, consequently, improving
predictive accuracy.

Furthermore, the importance of standardization cannot be overstated, especially
for predictors related to food groups. Standardization ensures that all
variables are on the same scale, crucial when dealing with diverse food supplies
in varying absolute quantities. Since some food groups may have low consumption
rates, standardization aids in providing a fair comparison between different
predictors, contributing to the overall effectiveness of the Lasso regression
model.

## Lasso Regression Model

```{r}
lasso.mod <- glmnet(x[train,], y[train], alpha = 1)
par(mar = c(6, 6, 6, 6))
plot(lasso.mod, main = "Lasso Regularization Path")
```

```{r}
# CV
set.seed(1)
cv.out <- cv.glmnet(x[train,], y[train], alpha = 1)
par(mar = c(6, 6, 6, 6))
plot(cv.out, main = "Lasso Prediction Error")
bestlam <- cv.out$lambda.min
lasso.pred <- predict(lasso.mod, s = bestlam, newx = x[test,])
```

```{r results = "hide"}
print(bestlam)
```

To fit the Lasso model, a ten-fold cross-validation (CV) approach is employed to
identify the optimal lambda value, which is determined to be 0.23. This lambda
value is subsequently utilized to construct the final model using the entire
dataset.

## Model Comparison

```{r results = "hide"}
# Simple LM MSE
mean((lm.pred - y.test)^2)
```

```{r results = "hide"}
# Lasso MSE
mean((lasso.pred - y.test)^2)
```

When comparing the prediction accuracy of the Lasso model with the simple linear
regression model, a substantial improvement is evident. The testing Mean Squared
Error (MSE) decreased significantly, dropping from 44.6 in the simple linear
regression model to 7.8 in the Lasso model. This noteworthy reduction
underscores the enhanced performance of the Lasso model, emphasizing its
superior predictive capabilities in capturing the underlying patterns within the
data.

# Results

## Final Model (Lasso)

```{r results = "hide"}
# Lasso
out <- glmnet(x, y, alpha = 1)
lasso.coef <- predict(out, type = "coefficients", s = bestlam)
lasso.coef
```

```{r results = "hide"}
# Convert the coefficient matrix to a regular matrix
lasso_coefs_matrix <- as.matrix(lasso.coef)

# Convert the matrix to a data frame
lasso_coefs_df <- as.data.frame(lasso_coefs_matrix)
lasso_coefs_df <- round(lasso_coefs_df, 4)

colnames(lasso_coefs_df) <- c("Coefficient")

kable(lasso_coefs_df)
```

```{r}
lasso_coefs_df <- lasso_coefs_df |>
  filter(Coefficient != 0)

kable(lasso_coefs_df)
```

## Model Interpretation

The table above presents the variables selected by the Lasso model along with
their corresponding coefficients. These coefficients indicate the strength and
direction of the relationship between each predictor and the obesity-attributed
death rate. Here is an interpretation of the selected variables:

**Intercept**: The intercept of 5.48% represents the predicted
obesity-attributed death rate when the supply of all food groups is
theoretically zero, and the country is classified as part of the High-income
group. It's essential to note that this scenario is not practically meaningful,
as complete absence of food supply is unrealistic. The intercept serves as a
baseline reference point for the model but should be interpreted with caution in
the context of food-related variables.

**Oilcrops**: The positive coefficient for Oilcrops indicates that a one-unit
increase in the supply of oil-rich crops, measured in kilocalories per day per
capita, is associated with an estimated 0.80% increase in the obesity-attributed
death rate, holding other variables constant. This relationship suggests that a
higher intake of oil-rich crops, such as certain seeds and nuts, may contribute
to a diet with a higher fat content, potentially influencing obesity-related
health issues.

**Fish and Seafood**: The negative coefficient for Fish and Seafood suggests
that a one-unit increase in the supply of fish and seafood is associated with an
estimated 1.72% decrease in the obesity-attributed death rate, holding other
variables constant. This aligns with the established health benefits of fish
consumption, as fish is a rich source of omega-3 fatty acids, which may
contribute to improved cardiovascular health and a potential reduction in
obesity-related risks.

**Sugar Sweeteners**: The positive coefficient for Sugar Sweeteners indicates
that a one-unit increase in the supply of sugar sweeteners is associated with an
estimated 0.35% increase in the obesity-attributed death rate, holding other
variables constant. This suggests that higher supplies of sugar sweeteners,
often found in processed foods and beverages, may contribute to an increased
risk of obesity-related health issues due to excessive sugar consumption.

**Meat Sheep and Goat**: The positive coefficient for Meat Sheep and Goat
suggests that a one-unit increase in the supply of sheep and goat meat is
associated with an estimated 0.0145% increase in the obesity-attributed death
rate, holding other variables constant. This relationship may be explained by
the higher saturated fat content in these meats, which, when consumed
excessively, may contribute to adverse health outcomes related to obesity.

**Meat Pig**: The negative coefficient for Meat Pig indicates that a one-unit
increase in the supply of pig meat is associated with an estimated 0.0032%
decrease in the obesity-attributed death rate, holding other variables constant.
This may be due to a potential protective effect, as pig meat is often leaner
compared to some other meats, and moderate consumption of lean protein may be
associated with better health outcomes.

**Meat Poultry**: The positive coefficient for Meat Poultry suggests that a
one-unit increase in the supply of poultry meat is associated with an estimated
0.0242% increase in the obesity-attributed death rate, holding other variables
constant. This relationship may be associated with the consumption of processed
poultry products, such as deli meats and deep-fried products, which can be high
in unhealthy fats and additives, potentially contributing to obesity-related
health issues.

**Meat Beef**: The negative coefficient for Meat Beef indicates that a one-unit
increase in the supply of beef is associated with an estimated 0.0154% decrease
in the obesity-attributed death rate, holding other variables constant. This
suggests a potential protective effect, possibly because lean beef can be part
of a balanced diet, providing essential nutrients without excessive unhealthy
fats.

**Fruit**: The positive coefficient for Fruit indicates that a one-unit increase
in the supply of fruits is associated with an estimated 0.0004% increase in the
obesity-attributed death rate, holding other variables constant. This
relationship may be explained by the fact that while fruits contain essential
vitamins and fiber, excessive consumption of certain fruits high in natural
sugars may contribute to caloric intake and, in turn, obesity-related risks.

**Pulses**: The negative coefficient for Pulses suggests that a one-unit
increase in the supply of pulses is associated with an estimated 0.0013%
decrease in the obesity-attributed death rate, holding other variables constant.
This potential protective effect may be attributed to the high fiber and plant
protein content in pulses, contributing to satiety and a balanced diet, reducing
the risk of obesity-related health issues.

**Cereals Other**: The negative coefficient for Cereals Other indicates that a
one-unit increase in the supply of other cereals is associated with an estimated
0.0021% decrease in the obesity-attributed death rate, holding other variables
constant. This relationship may be associated with the inclusion of whole grains
in this category, known for their health benefits, including promoting satiety
and contributing to better weight management.

**Wheat**: The positive coefficient for Wheat indicates that a one-unit increase
in the supply of wheat is associated with an estimated 0.0063% increase in the
obesity-attributed death rate, holding other variables constant. This suggests a
potential increase in obesity-related health risks with higher wheat supply,
possibly due to the consumption of refined wheat products lacking the
nutritional benefits of whole grains.

**Income Groups**:

**Lower-middle-income countries**: The positive coefficient for
Lower-middle-income countries indicates that being classified as part of this
income group is associated with an estimated 0.0289% increase in the
obesity-attributed death rate, holding other variables constant. This
relationship may be influenced by various factors, including changes in dietary
patterns, lifestyle, and healthcare infrastructure that contribute to
obesity-related health risks.

**Upper-middle-income countries**: The significant positive coefficient for
Upper-middle-income countries suggests a more substantial impact, with an
estimated 1.8862% increase in the obesity-attributed death rate, holding other
variables constant. This may reflect a complex interplay of dietary, lifestyle,
and socioeconomic factors in upper-middle-income countries contributing to
heightened obesity-related health risks.

## Implications

The findings from this analysis hold substantial implications for policymakers
and consumers, providing actionable insights to combat the escalating issue of
obesity-related deaths. For policymakers, the identified significant predictors
offer a roadmap for targeted intervention strategies. The positive coefficient
for Lower-middle-income countries, for instance, signals a need for tailored
public health initiatives in these economic contexts. Policymakers can leverage
this information to design impactful public health campaigns, promoting the
consumption of healthier food choices, such as Fish and Seafood, to mitigate the
adverse effects of obesity. The insights into the economic considerations, as
highlighted by the Income Group coefficients, allow policymakers to allocate
resources efficiently, considering economic disparities in addressing
obesity-related health risks. Additionally, consumers can benefit from this
analysis by making informed dietary choices. Understanding the impact of
specific food groups, such as Sugar Sweeteners, encourages moderation in the
consumption of processed foods, while the negative coefficient for Fish and
Seafood suggests substituting red meat with fish for potential health benefits.
Moreover, the nuanced effects of different meat types provide consumers with
insights for achieving a balanced diet. By incorporating whole grains and
fiber-rich foods, consumers can contribute to better weight management. The
awareness of economic disparities influencing obesity-related death rates
prompts consumers to consider the broader socioeconomic context and advocate for
systemic changes. In summary, these findings empower both policymakers and
consumers with evidence-based information to address obesity-related deaths
comprehensively.

# Conclusion

This project explored the global impact of obesity-related deaths, focusing on
the interplay of food supply, economic status, and mortality rates. Key findings
highlighted the severity of obesity-related health crises and the economic
burden. The analysis identified significant factors using a dual-stage
methodology, revealing nuanced associations.

Economic status, represented by income groups, showed distinct links.
Lower-middle-income countries were associated with increased death rates,
emphasizing the need for targeted interventions. Upper-middle-income countries
exhibited a significant rise in mortality rates, warranting further
investigation.

Food groups played a crucial role, with positive associations found for oil-rich
crops and sugar sweeteners, indicating potential health risks. Negative
associations with fish and seafood suggested protective effects, guiding
healthier dietary choices.

The Lasso regression model refined variable selection, emphasizing the impact of
specific food groups and economic factors. These insights hold practical
implications for policymakers in crafting targeted interventions and resource
allocation. For consumers, the findings empower informed dietary choices to
manage weight and enhance overall health.

In conclusion, this project contributes evidence-based insights to address the
complex challenge of obesity-related deaths globally, providing a foundation for
comprehensive strategies.

## Limitations

1)  supply, not consumption. assuming consumption is proportional to supply.

# Citations

[1] <https://www.cdc.gov/obesity/>

[2] <https://ourworldindata.org/obesity>
